---
title: "P8105 Hw 1 glj2117"
output: 
  html_document:
  toc: true

---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

# Problem 1

**Penguins dataset summary.**


```{r}
data("penguins", package = "palmerpenguins")
head(penguins)

```
The variables of note included in this dataset are species, island of origin, bill and flipper length, bill depth, body mass, and sex. The dataset is `r nrow(penguins)` rows long and `r ncol(penguins)` variables wide. 
The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm. The maximum bill length is `r max(pensguins$bill_length_mm)` and the minimum is `r min(penguins$bill_length_mm)`. The maximum body mass is `r max(penguins$body_mass_g)` and the mean body mass is `r mean(penguins$body_mass_g)`.

```{r}

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("flipper_vs_bill_length.pdf")

```
# Problem 2
This sample data data frame is comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels” ("high", "med", "low")

```{r}

problem2_df = tibble(
  var_norm = rnorm(10),
  var_logical = var_norm > 0,
  var_characater = c("12 feet", "5 feet", "3 feet", "4 feet", "11 feet", "8 feet", "7 feet", "2 feet", "9 feet", "15 feet"),
  var_factor = factor(c("high", "med", "low", "low", "high", "med", "med", "low", "high", "high"))
)
```

```{r chunk_histogram}

```

```{r chunk_learning assessment}
  

```

